\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-12-30 archive.tex
more writing

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ec2},
each with 15 input/output examples.  In solving these tasks, the
system composed 16 new DSL components, and discovered multiple
higher-order functions, such as \code{filter} ($f_0$ in
Figure~\ref{initialExampleDSL}).  Each round of memory consolidation
built on components discovered in earlier sleep cycles --- for example
the agent first learns \code{filter}, uses filter to learn to take the
maximum element of a list, then uses that routine to learn a new
component for extracting the $n^{\text{th}}$ largest element of a
list, which it finally uses to learn how to sort a list of numbers
(Figure~\ref{initialExampleDSL}).  This incremental, modular learning
of deep hierarchies of DSL components is characteristic of \system
because of the alternation between code writing and code refactoring.

@@

\subsection{Programs that make plans and take actions}

@@

\begin{figure}
  \includegraphics[width = 0.3\textwidth]{figures/logo16.png}
  \includegraphics[width = 0.3\textwidth]{figures/tower9.png}
  \caption{LOGO graphics tasks. The agent must learn to write programs that draw these 90 pictures.}\label{everyLogo}
\end{figure}
\begin{figure}
\centering  \begin{tabular}{cc}
    \toprule
    Parametric drawing routines&Higher-order drawing routine\\\midrule
    \begin{tabular}{rl}
      Semicircle:& \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_24.png}}\\
      Circles:
      &\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_23.png}}\\
      Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_25.png}}\\
      Greek Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_10.png}}\\
      S-Curves:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_11.png}}        \\
        Polygons \& Stars:
&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_12.png}}
    \end{tabular}
    &
    \begin{tabular}{c}
      Repeat and rotate subprogram:\\
          \includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_18.png}
      \end{tabular}

@@

Inspired by the classic AI `copy task' --- where an agent must look at
an image of a tower made of toy blocks and re-create the
tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
(Figure~\ref{towerTasks}).  Here the agent observes both an image of a
tower and the locations of each of its blocks, and must write a
program that plans how a simulated hand would build the tower.
These towers are built from Lego-style blocks
that snap together on a discrete grid (e.g., we do not model gravity
and discretize the coordinates of the blocks).

The system starts out with four domain-specific primitives:
\code{left} and \code{right}, which take an integer $n$ as input and
move the hand left/right $n$ Lego grid spaces; along with
\code{horizontal} and \code{vertical}, which put down
horizontal/vertical Lego blocks.  As with LOGO graphics we also
include our two standard planning primitives (\code{for} and
\code{get/set}).
Starting with this basis
the system can only solve
12.5\% of the copy tasks.
With successive wake/sleep cycles
we assemble a DSL containing
parametric option-like subroutines (Figure~\ref{towerPrimitives}),
allowing the agent to solve 90\% of the tasks.
\begin{figure}
  \caption{Learned subroutines for building towers out of Lego-style blocks when trained on tasks in Figure~\ref{towerTasks}. These subroutines act like parametric options~\cite{options},
    giving higher-level building blocks that the agent can use to plan.}
\end{figure}

\subsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
We apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 %% symbolic regression
problems, each either a polynomial or rational function.  The recognition model is a
convnet that observes an image of the target function's
graph (Fig.~\ref{functions}) --- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the convnet can  look at a graph and predict
what kind of function best explains it.
A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to enumerate
 programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
\begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.1cm}\centering
  \begin{minipage}[c]{0.3\columnwidth}%\vspace{-0.3cm}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/48.png}\\
  \includegraphics[width = \functionSize]{figures/functions/102.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
  \vspace{1pt}%\includegraphics[width = \functionSize]{figures/functions/181.png}
  %\includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
    \includegraphics[width = \functionSize]{figures/functions/149.png}
  \end{minipage}
  \begin{minipage}[c]{0.69\columnwidth}    
    \caption{Recognition model input for symbolic regression. DSL learns subroutines for polynomials (top rows) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (left) and predict which learned subroutines best explain the observation.}\label{functions}\vspace{-1cm}
        \end{minipage}
\end{figure}
We learn a DSL containing 13 new functions,
mainly templates for polynomials of different orders or ratios of polynomials.
The model also learns to find programs minimizing the number of continuous parameters ---
learning to represent linear functions with 
\code{(* real (+ x real))}, which has two continuous parameters, and represents quartic functions using $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing:
both the generative model's bias toward shorter programs,
and the likelihood model's BIC penalty.

@@

After running for 93 hours on 64 CPUs, our
algorithm solves these 22 exercises, along the way assembling a DSL
with a modern repertoire of
functional programming idioms and subroutines, including \code{map},
\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
and  arithmetic operations like 
building lists of natural numbers between an interval (see Figure~\ref{learningFromScratch}).
We did not use the recognition model for this experiment:
a bottom-up pattern recognizer is of little use
for acquiring this abstract knowledge
from less than two dozen problems.

\begin{figure*}  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{cc}\toprule
    \normalsize \pop{Programs} \& Tasks&\normalsize \popp{DSL}\\\midrule

We believe that program learners should \emph{not}
start from scratch,
but instead should start from
a rich, domain-agnostic
basis like those embodied in the standard libraries of modern  languages.
What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
and can in principle recover many of the amenities of modern programming systems,
provided it is given enough computational power and a suitable
spectrum of tasks.% programming exercises.

\subsection{Learning Generative Models}

We apply \system to learning generative models for images and text (Fig.~\ref{geomCompiled}-\ref{regularExpressions}).
For images, we learn programs controlling a simulated ``pen,''
and the task is to look at an image and explain it in terms of a graphics program.
For text, we learn probabilistic regular expressions -- a simple probabilistic program for which inference is always tractable -- and the task is to
infer a regex from a collection of strings.

\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
We evaluate  on held-out testing tasks,
measuring how many
tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
Prior to any learning,
the system cannot find solutions for most of the tasks,
and those it does solve take a long time;
with more wake/sleep iterations,
we converge upon DSLs and recognition models 
more closely matching the domain.

\begin{figure}
  \includegraphics[width = 5cm]{figures/logoLearningCurve.png}
  \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}
  \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}
  \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}
  \end{figure}

\begin{figure*}\centering
  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps} \qquad
  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}\qquad
  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
\vspace{-0.4cm}  \caption{Learning curves for \system both with (\orange{in orange}) and without
    (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved w/ 10m timeout. Dashed lines: Average solve time, averaged only over tasks that are solved.}\label{learningCurves}\vspace{-0.5cm}
\end{figure*}

%\paragraph{Outlook.}
We contribute an algorithm, \systemEnding, that learns to program by
bootstrapping a DSL with new domain-specific primitives that the algorithm
itself discovers, together with a neural recognition model that learns to
efficiently deploy the DSL on new tasks. %% We believe this integration of top-down
%% symbolic representations and bottom-up neural nets --- both of them learned
%% --- can help make program induction systems more generally useful for AI. 
%Many
%directions remain open.
%\paragraph{Future.}
Two immediate future goals are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in some
domains over the generic methods used here.
Another direction is DSL meta-learning: can we find a
\emph{single} universal primitive set that could bootstrap DSLs for
new domains, including the domains considered here,  but also many others?

@@

     \State \textbf{function} recognitionSample$'(Q, x, \text{parent}, \text{argumentIndex}, \mathcal{D}, \mathcal{E}, \tau)$:
     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, parent $\in \mathcal{D}\cup\left\{\text{start},\text{var} \right\}$, argumentIndex $\in \mathbb{N}$, environment $\mathcal{E}$, type $\tau$
     \State \textbf{Output:} a program whose type unifies with $\tau$
     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
     \State var $\gets$ an unused variable name
     \State body $\sim$ recognitionSample$'(Q,x,\text{parent},\text{argumentIndex},\mathcal{D},\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$%\Comment{Recursively sample function body}
     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
     %\Endif
     \Else\Comment{Build an application to give something w/ type $\tau$}
     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
       Q_{\text{parent},e,\text{argumentIndex}}(x)&\text{ if }e\in \mathcal{D}\\
Q_{\text{parent},\text{var},\text{argumentIndex}}(x)/|\text{variables}|&\text{ if }e\in \mathcal{E}
       \end{cases}$
     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
     \State newParent$\gets\begin{cases}
     e &\text{ if }e\in \mathcal{D}\\
     \text{var}&\text{ if }e\in \mathcal{E}\end{cases}$
     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
     %   \State unify$(\tau,\beta)$
     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
     \State $a_k\sim\text{recognitionSample}'(Q,x,\text{newParent},k,\mathcal{D},\mathcal{E},\alpha_k)$
     \EndFor
     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
     \EndIf
     %% \Statex
     %% \Statex\textbf{where:}
     %% \State yield$(\tau) = \begin{cases}
     %%   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
     %%   \tau   &\text{ otherwise.}
     %% \end{cases}$ \Comment{Final return type of $\tau$}
     %% \State  args$(\tau) = \begin{cases}
     %%   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
     %%   []   &\text{ otherwise.}
     %% \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
   \end{algorithmic}
 \end{algorithm}

 \textbf{Symmetry breaking analysis.}

 \begin{theorem}
   Let $\mu(\cdot )$ be a distribution over tasks and let $Q^*(\cdot |\cdot )$ be a task-conditional distribution over programs satisfying
   $$
   Q^* = \argmax_Q\;\expect_\mu\left[\max_{\substack{p\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}}\log  Q(p|x) \right]
   $$
   where $(\mathcal{D},\theta)$ is a generative model over programs. Pick a task $x$ where $\mu(x) > 0$.
Partition $\Lambda$ into expressions that are observationally equivalent under $x$:
   $$
\Lambda = \bigcup_i \mathcal{E}^x_i\text{ where for any }p_1\in \mathcal{E}^x_i\text{ and }p_2\in \mathcal{E}^x_j\text{: }\probability[x|p_1] = \probability[x|p_2] \iff i = j
$$
Then there exists an equivalence class $\mathcal{E}^x_i$
that gets all the probability mass of $Q^*$ -- e.g.,
$Q^*(p|x) = 0$ whenever $p\not\in \mathcal{E}^x_i$ --
and there exists a program in that equivalence class
which gets all of the probability mass assigned by $Q^*(\cdot |x)$ -- e.g.,  there is a $p\in \mathcal{E}^x_i$
such that $Q^*(p|x) = 1$ -- and that program maximizes $\probability[\cdot |x,\mathcal{D},\theta ]$.
 \end{theorem}
 \begin{proof}
   We proceed by defining the set of ``best programs'' -- programs maximizing the posterior $\probability[\cdot |x,\mathcal{D},\theta]$ -- and then showing that a best program satisfies $Q^*(p|x) = 1$.
   Define the set of best programs $\mathcal{B}_x$ for the task $x$ by
   $$
\mathcal{B}_x = \left\{p\;|\;\probability[p|x,\mathcal{D},\theta] = \max_{p'\in \Lambda}\probability[p'|x,\mathcal{D},\theta] \right\}
$$
For convenience define
$$
f(Q) = \expect_\mu\left[\max_{p\in \mathcal{B}_x}\log Q(p|x) \right]
$$
and observe that $Q^* = \argmax_Q f(Q)$.

@@

 Because the equivalence classes $\left\{\mathcal{E}_i^x \right\}$
 form a partition of $\Lambda$
 we know that $p^*$ is a member of exactly one equivalence class; call it $\mathcal{E}_*^x$.
 This accomplishes one of our goals:
 we have already shown that $Q^*(p^*|x) = 1$.

 Let $q\in \mathcal{E}_i^x\not=\mathcal{E}_*^x$.
 Then because the equivalence classes form a partition we know that
 $q\not= p^*$
 and so $Q^*(q|x) = 0$,
 which is what was to be shown:
 \emph{any}
 program not in $\mathcal{E}_*^x$
 gets no probability mass.

 \end{proof}

 Consider an agent tasked with writing programs built from addition and the constants zero and one.
A bigram parameterization of $Q$ allows
it to represent the fact that it should never add zero ($Q_{\code{+},\code{0},0} = Q_{\code{+},\code{0},1} = 0$)
%or never multiply by  1 ($Q_{\code{*},\code{1},0} = Q_{\code{*},\code{1},1} = 0$), or
or that addition should always associate to the right
($Q_{\code{+},\code{+},0} = 0$).
The $\mathcal{L}^{\text{MAP}}$ training objective encourages
learning these canonical forms.
Consider two recognition models, $Q_1$ and $Q_2$,
and two programs in frontier $\mathcal{F}_x$,
$p_1=\code{(+ (+ 1 1) 1)}$ and $p_2=\code{(+ 1 (+ 1 1))}$,
where
\begin{align*}
  Q_1(p_1|x) = \frac{\epsilon}{2}&\qquad Q_1(p_2|x) = \frac{\epsilon}{2}\\
  Q_2(p_1|x) = 0&\qquad  Q_2(p_2|x) = \epsilon
\end{align*}
i.e., $Q_2$ breaks a symmetry by forcing right associative addition,
but $Q_1$ does not, instead splitting its probability mass equally between $p_1$ and $p_2$.
Now because $\probability[p_1|\mathcal{D},\theta] = \probability[p_2|\mathcal{D},\theta]$
(Algorithm~\ref{sampleProgram}),
we have
\begin{align*}
  \mathcal{L}^{\text{posterior}}_{\text{real}}(Q_1)& = \frac{\probability[p_1|\mathcal{D},\theta]\log \frac{\epsilon}{2} + \probability[p_2|\mathcal{D},\theta]\log \frac{\epsilon}{2}}{\probability[p_1|\mathcal{D},\theta] + \probability[p_2|\mathcal{D},\theta]} = \log \frac{\epsilon}{2}\\
  \mathcal{L}^{\text{posterior}}_{\text{real}}(Q_2)& = \frac{\probability[p_1|\mathcal{D},\theta]\log 0 + \probability[p_2|\mathcal{D},\theta]\log \epsilon}{\probability[p_1|\mathcal{D},\theta] + \probability[p_2|\mathcal{D},\theta]} = +\infty\\
  \mathcal{L}^{\text{MAP}}_{\text{real}}(Q_1)& = \log Q_1(p_1)           = \log Q_1(p_2)   = \log \frac{\epsilon}{2}\\
  \mathcal{L}^{\text{MAP}}_{\text{real}}(Q_2)& = \log Q_2(p_2) \phantom{ = \log Q_1(p_2) } = \log \epsilon\\
\end{align*}
So $\mathcal{L}^{\text{MAP}}$ prefers $Q_2$ (the symmetry breaking
recognition model), while $\mathcal{L}^{\text{posterior}}$ reverses
this preference.

>>>>

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ec2},
each with 15 input/output examples.  In solving these tasks, the
system composed 16 new DSL components, and discovered multiple
higher-order functions, such as \code{filter} ($f_0$ in
Figure~\ref{initialExampleDSL}).  Each round of memory consolidation
built on components discovered in earlier sleep cycles --- for example
the agent first learns \code{filter}, uses filter to learn to take the
maximum element of a list, then uses that routine to learn a new
component for extracting the $n^{\text{th}}$ largest element of a
list, which it finally uses to learn how to sort a list of numbers
(Figure~\ref{initialExampleDSL}).  This incremental, modular learning
of deep hierarchies of DSL components occurs
because of the alternation between code writing and code refactoring.

@@

\subsection{Programs from visual input}
We consider three domains where the agent must infer a program from an
image (Figure~\ref{visualSpecs}).
\subsubsection{Programs that make plans and take actions}

@@

\begin{figure}\centering
  \includegraphics[width = 0.3\textwidth]{figures/logo16.png}
  \includegraphics[width = 0.3\textwidth]{figures/tower9.png}
  \includegraphics[width = 0.3\textwidth]{figures/sr16.png}
  \caption{Three domains where the agent infers a program from visual input. \textbf{Left}: 16 (out of 145) LOGO graphics tasks. Agent writes a program controlling a `pen' that draws the target picture. \textbf{Middle}: 9 (out of 112) tower building tasks. Agent writes a program controlling a `hand' that builds the target tower. \textbf{Right}: 16 (out of 200) symbolic regression tasks. Agent writes a program containing continuous real numbers that fits the points along the curve.}\label{visualSpecs}
\end{figure}
\begin{figure}
\centering  \begin{tabular}{cc}
    \toprule
    Parametric drawing routines&Higher-order drawing routine\\\midrule
    \begin{tabular}{rl}
      Semicircle:& \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_24.png}}\\
      Circles:
      &\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_23.png}}\\
      Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_25.png}}\\
      Greek Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_10.png}}\\
      S-Curves:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_11.png}}        \\
        Polygons \& Stars:
&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_12.png}}
    \end{tabular}
    &
    \begin{tabular}{c}
      Rotational Symmetry:\\
          \includegraphics[width = 0.35\textwidth]{figures/rotationalmontage.png}
      \end{tabular}

@@

Inspired by the classic AI `copy task' --- where an agent must look at
an image of a tower made of toy blocks and re-create the
tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
(Figure~\ref{towerTasks}).  Here the agent observes both an image of a
tower and the locations of each of its blocks, and must write a
program that plans how a simulated hand would build the tower.
These towers are built from Lego-style blocks
that snap together on a discrete grid.%%  (e.g., we do not model gravity
%% and discretize the coordinates of the 
%% blocks).

The system starts out with four domain-specific primitives:
\code{left} and \code{right}, which take an integer $n$ as input and
move the hand left/right $n$ Lego grid spaces; along with
\code{horizontal} and \code{vertical}, which put down
horizontal/vertical Lego blocks.  As with LOGO graphics we also
include our two standard planning primitives (\code{for} and
\code{get/set}).
Starting with this basis
the system can only solve
12.5\% of the copy tasks.
\begin{figure}
  \begin{tabular}{cc}
    \toprule Example learned DSL components&Samples from DSL\\\midrule
    \begin{tabular}{rl}
      Brickwall&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bricks.png}
      \end{tabular}\\
      Bridge&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bridge.png}
      \end{tabular}\\
      Staircase&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_staircase.png}
        \end{tabular}
    \end{tabular}&
    \begin{tabular}{l}
      \includegraphics[width = 5cm]{figures/tower/dreams.png}
      \end{tabular}
    \\\bottomrule 
    \end{tabular}
  \caption{Learned subroutines for building towers out of Lego-style blocks. These subroutines act like parametric options~\cite{options},
    giving higher-level building blocks that the agent can use to plan.}
\end{figure}

\subsubsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
Here, the agent observes points along the curve of a function, and
must write a program that fits those points.  We initially equip our
learner with addition, multiplication, and division, and task it with
solving 200 symbolic regression problems, each either a polynomial or
rational function.  The recognition model is a convnet that observes
an image of the target function's graph (Fig.~\ref{visualSpecs},
rightmost) --- visually, different kinds of polynomials and rational
functions produce different kinds of graphs, and so the convnet can
look at a graph and predict what kind of function best explains it.  A
key difficulty, however, is that these problems are best solved with
programs containing real numbers.  Our solution to this difficulty is
to enumerate programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.  We
define the likelihood model, $\probability[x|p]$, by assuming a
Gaussian noise model for the input/output examples, and penalize the
use of real-valued parameters using the
BIC~\cite{Bishop:2006:PRM:1162264}.

We learn a DSL containing 13 new functions, mainly templates for
different pieces of polynomials or ratios of polynomials.  The model
also learns to find programs minimizing the number of continuous
parameters --- for example, learning to represent linear functions
with \code{(* real (+ x real))}.  This phenomenon arises from our
Bayesian framing: both the generative model's bias toward shorter
programs, and the likelihood model's BIC penalty.

h
\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
We evaluate  on held-out testing tasks,
measuring how many
tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
Prior to any learning,
the system cannot find solutions for most of the tasks,
and those it does solve take a long time;
with more wake/sleep iterations,
we converge upon DSLs and recognition models 
more closely matching the domain.

\begin{figure}
  \begin{tabular}{ccc}
    \includegraphics[width = 5cm]{figures/logoLearningCurve.png}&
    \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\
    \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/rationalLearningCurve_batched.png}&
  \end{tabular}
  \caption{}\label{learningCurves}
\end{figure}

@@

After running for 93 hours on 64 CPUs, our
algorithm solves these 22 exercises, along the way assembling a DSL
with a modern repertoire of
functional programming idioms and subroutines, including \code{map},
\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
and  arithmetic operations like 
building lists of natural numbers between an interval (see  Appendix~\ref{appendixMcCarthy}).

We believe that program learners should \emph{not}
start from scratch,
but instead should start from
a rich, domain-agnostic
basis like those embodied in the standard libraries of modern  languages.
What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
and can in principle recover many of the amenities of modern programming systems,
provided it is given enough computational power and a suitable
spectrum of tasks.

\subsection{Program Induction as part of the generic AI toolkit}

Our aim with \system is both to show how learning can enable better
program synthesis, and to show how better program synthesis can be
more broadly useful for AI. How could a program induction approach to
AI become broadly useful across a diverse spectrum of AI problems ---
planning, natural language understanding, machine vision, causal
inference, and more?

@@

     \State \textbf{function} recognitionSample$'(Q, x, \text{parent}, \text{argumentIndex}, \mathcal{D}, \mathcal{E}, \tau)$:
     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, parent $\in \mathcal{D}\cup\left\{\text{start},\text{var} \right\}$, argumentIndex $\in \mathbb{N}$, environment $\mathcal{E}$, type $\tau$
     \State \textbf{Output:} a program whose type unifies with $\tau$
     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
     \State var $\gets$ an unused variable name
     \State body $\sim$ recognitionSample$'(Q,x,\text{parent},\text{argumentIndex},\mathcal{D},\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$%\Comment{Recursively sample function body}
     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
     %\Endif
     \Else\Comment{Build an application to give something w/ type $\tau$}
     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
       Q_{\text{parent},e,\text{argumentIndex}}(x)&\text{ if }e\in \mathcal{D}\\
Q_{\text{parent},\text{var},\text{argumentIndex}}(x)/|\text{variables}|&\text{ if }e\in \mathcal{E}
       \end{cases}$
     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
     \State newParent$\gets\begin{cases}
     e &\text{ if }e\in \mathcal{D}\\
     \text{var}&\text{ if }e\in \mathcal{E}\end{cases}$
     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
     %   \State unify$(\tau,\beta)$
     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
     \State $a_k\sim\text{recognitionSample}'(Q,x,\text{newParent},k,\mathcal{D},\mathcal{E},\alpha_k)$
     \EndFor
     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
     \EndIf
   \end{algorithmic}
 \end{algorithm}

 \noindent \textbf{Symmetry breaking.}  Why does the combination of
 $\mathcal{L}^{\text{MAP}}$ and the bigram parameterization lead to
 symmetry breaking?  The reason is twofold: (1) the objective
 $\mathcal{L}^{\text{MAP}}$ prefers symmetry breaking recognition
 models; and (2) the bigram parameterization permits certain kinds of
 symmetry breaking.
 To sharpen these intuitions,
 we prove (Theorem~\ref{symmetricproof})
 that any global optimizer of $\mathcal{L}^{\text{MAP}}$
 breaks symmetries,
 and then give a concrete worked out example
 contrasting the behavior of $\mathcal{L}^{\text{MAP}}$ and $\mathcal{L}^{\text{posterior}}$.

 \pagebreak\begin{theorem}\label{symmetricproof}
   Let $\mu(\cdot )$ be a distribution over tasks and let $Q^*(\cdot |\cdot )$ be a task-conditional distribution over programs satisfying
   $$
   Q^* = \argmax_Q\;\expect_\mu\left[\max_{\substack{p\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}}\log  Q(p|x) \right]
   $$
   where $(\mathcal{D},\theta)$ is a generative model over programs. Pick a task $x$ where $\mu(x) > 0$.
Partition $\Lambda$ into expressions that are observationally equivalent under $x$:
   $$
\Lambda = \bigcup_i \mathcal{E}^x_i\text{ where for any }p_1\in \mathcal{E}^x_i\text{ and }p_2\in \mathcal{E}^x_j\text{: }\probability[x|p_1] = \probability[x|p_2] \iff i = j
$$
Then there exists an equivalence class $\mathcal{E}^x_i$
that gets all the probability mass of $Q^*$ -- e.g.,
$Q^*(p|x) = 0$ whenever $p\not\in \mathcal{E}^x_i$ --
and there exists a program in that equivalence class
which gets all of the probability mass assigned by $Q^*(\cdot |x)$ -- e.g.,  there is a $p\in \mathcal{E}^x_i$
such that $Q^*(p|x) = 1$ -- and that program maximizes $\probability[\cdot |x,\mathcal{D},\theta ]$.
 \end{theorem}
 \begin{proof}
   We proceed by defining the set of ``best programs'' -- programs maximizing the posterior $\probability[\cdot |x,\mathcal{D},\theta]$ -- and then showing that a best program satisfies $Q^*(p|x) = 1$.
   Define the set of best programs $\mathcal{B}_x$ for the task $x$ by
   $$
\mathcal{B}_x = \left\{p\;|\;\probability[p|x,\mathcal{D},\theta] = \max_{p'\in \Lambda}\probability[p'|x,\mathcal{D},\theta] \right\}
$$
For convenience define
$$
f(Q) = \expect_\mu\left[\max_{p\in \mathcal{B}_x}\log Q(p|x) \right]
$$
and observe that $Q^* = \argmax_Q f(Q)$.

@@

 Because the equivalence classes $\left\{\mathcal{E}_i^x \right\}$
 form a partition of $\Lambda$
 we know that $p^*$ is a member of exactly one equivalence class; call it $\mathcal{E}_i^x$.
 Let $q\in \mathcal{E}_j^x\not=\mathcal{E}_i^x$.
 Then because the equivalence classes form a partition we know that
 $q\not= p^*$
 and so $Q^*(q|x) = 0$,
 which was our first goal:
 \emph{any}
 program not in $\mathcal{E}_i^x$
 gets no probability mass.

 Our second goal --- that there is a member of $\mathcal{E}_i^x$ which gets all the probability mass assigned by $Q^*(\cdot |x)$ --- is immediate from $Q^*(p^*|x) = 1$.

 Our final goal --- that $p^*$ maximizes $\probability[\cdot |x,\mathcal{D},\theta]$
 --- follows from the fact that $p^*\in \mathcal{B}_x$.
 \end{proof}

 Notice that Theorem~\ref{symmetricproof} makes no guarantees as to the
 cross-task systematicity of the symmetry breaking; for example, an
 optimal recognition model could associate addition to the right for
 one task and associate addition to the left on another task. \emph{Systematic} breaking of symmetries
 must arise only as a consequence as
 the network architecture (i.e., it is more parsimonious to break symmetries the same way for every task than it is to break them differently for each task).
